\chapter*{Conclusion}
\label{ch:conclusion}

Details about ANNs -- the second generation of neural networks, and details about SNNs -- third generation were presented. The state-of-the-art activation functions, spiking neurons, and learning approaches as well. In following part, we discussed multiple simulation tools, which are useful for modeling NNs. In the end of the theoretical part, new and promising neuromorphic hardware was presented.

In the other part of the thesis, used dataset was described together with its processing and training neural models. Tensorflow and NengoDL libraries were used or the experiment. Four models in total were implemented, 1D~Convolutional ANN, LSTM ANN, transformed 1D~Convolutional SNN from ANN, ans SNN with new LMU neurons. Models were trained on two differently processed datasets. At first after signal averaging with sliding window and then after randomly picked samples averaging. The latter proved as more effective. Best performing models were convolutional neural networks.

It would be interesting to use real neuromorphic hardware as a next step of this research. More complex spiking networks with bigger dataset and more demanding parameters could be tested. They should outperform SNNs used here on Von Neumann architecture. All codes and output files are available here \cite{thesis-code}.


\chapter*{List of abbreviations}

\begin{itemize}
	\item NN -- Neural Network
	\item ANN -- Artificial Neural Network
	\item SNN -- Spiking Neural Network
	\item DNN -- Deep Neural Network
	\item RNN -- Recurrent Neural Network
	\item MLP -- Multi Layered Perceptron
	\item LSTM -- Long Short-Term Memory
	\item GRU -- Gated Recurrent Networks
	\item SiLU -- Sigmoid-weighted Linear Unit
	\item STDP -- Spike Time-Dependent Plasticity
	\item LTP -- Long-Term Potentiation
	\item LTD -- Long-Term Depression
	\item AI -- Artificial Intelligence
	\item ML -- Machine Learning
	\item FAIR -- Facebook AI Research lab
	\item NEF -- Neural Engineering Framework
	\item HBP -- Human Brain Project
	\item EEG -- Electroencephalography
	\item LIF -- Leaky Integrate-and-Fire
	\item LMU -- Legendre Memory Unit
	\item ReLU -- Rectified Linear Unit
	\item NEST -- The Network Simulation Tool
	\item NMHW -- Neuromorphic Hardware
	\item PV1 -- Pre-process version 1
	\item PV2 -- Pre-process version 2
\end{itemize}
